---
title: "Week3"
author: "Marcelo Szilagyi"
date: "11/16/2016"
output: html_document
---

```{r setup, include = FALSE, echo = FALSE, cache = TRUE, results='hide'}
knitr::opts_chunk$set(echo = TRUE)
suppressMessages(suppressWarnings(
  sapply(listpackages, function (x) library(x,character.only = T))
  ))
```


Pre work: join all english datasets in one big corpus of text, but keep one reference
```{r}
englishfiles = list.files(recursive = TRUE, pattern = glob2rx('*en_*.txt'))
filenames = str_extract(englishfiles,"(?<=.)[^.]+(?=.txt)")

for (i in seq_along(englishfiles)) {
  assign(filenames[i], readLines(con = englishfiles[i],warn = FALSE))
 }

# create tibbles
twitter = as_tibble(twitter);blogs = as_tibble(blogs);news = as_tibble(news)

# add row number and origin for future use
twitter = twitter %>% mutate(linenumber = row_number(), origin = "t")
blogs = blogs %>% mutate(linenumber = row_number(), origin = "b")
news = news %>% mutate(linenumber = row_number(), origin = "n")

# join all tibbles in one big tibble
corpo = twitter
rbind(corpo,news,blogs)

```

Pre work: remove simbols (but not [Diacritics](https://en.wikipedia.org/wiki/Diacritic )
```{r}

```


## Exploratory Data Analysis


Tasks to accomplish

Questions to consider

1) Some words are more frequent than others - what are the distributions of word frequencies?
```{r}

```

2) What are the frequencies of 2-grams and 3-grams in the dataset?
```{r}

```


3) How many unique words do you need in a frequency sorted dictionary to cover 50% of all word instances in the language? 90%?
```{r}

```

4) How do you evaluate how many of the words come from foreign languages?
```{r}

```

5) Can you think of a way to increase the coverage -- identifying words that may not be in the corpora or using a smaller number of words
in the dictionary to cover the same number of phrases?
```{r}

```


Exploratory analysis - perform a thorough exploratory analysis of the data, understanding the distribution of words and relationship between the words in the corpora.
```{r}

```

Understand frequencies of words and word pairs - build figures and tables to understand variation in the frequencies of words and word pairs in the data.
```{r}

```



## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
